### 2008-10-10 trunk/Papers/rmrs-08/conclusion.tex
final?

In this paper, we have motivated and defined \rmrs---a semantic
framework that has been used to represent, compare, and combine
semantic information computed from deep and shallow parsers.  \rmrs\
is designed to be maximally flexible on the type of semantic
information that can be left underspecified, so that the semantic
output of a shallow parser needn't over-determine or under-determine
the semantics that can be extracted from the shallow syntactic
analysis.  The key contribution of this paper was to lay the formal
foundations for a formalism that is emerging as a standard in robust
semantic processing.

Although we have not directly provided new tools for modelling or
processing language, we believe that a cleanly defined model theory
for \rmrs\ is a crucial prerequisite for the future development of
such tools.  We hope that future research will build upon this paper
to develop efficient algorithms and implementations for solving \rmrs
s, performing inferences that enrich \rmrs s from shallow analyses
with deeper information, and checking consistency of \rmrs s that were
obtained from different parsers.

>>>>

In this paper, we motivated and defined \rmrs---a semantic framework
that has been used to represent, compare, and combine semantic
information computed from deep and shallow parsers.  \rmrs\ is
designed to be maximally flexible on the type of semantic information
that can be left underspecified, so that the semantic output of a
shallow parser needn't over-determine or under-determine the semantics
that can be extracted from the shallow syntactic analysis.  Our key
contribution was to lay the formal foundations for a formalism that is
emerging as a standard in robust semantic processing.

Although we have not directly provided new tools for modelling or
processing language, we believe that a cleanly defined model theory
for \rmrs\ is a crucial prerequisite for the future development of
such tools; this strategy was highly successful for dominance
constraints \cite{Althaus_etal:JoA}. We hope that future research will
build upon this paper to develop efficient algorithms and
implementations for solving \rmrs s, performing inferences that enrich
\rmrs s from shallow analyses with deeper information, and checking
consistency of \rmrs s that were obtained from different parsers.

### 2008-10-10 trunk/Papers/rmrs-08/entailment.tex
final?

The definitions from Section~\ref{sec:rmrs} serve an important
theoretical purpose in that they provide a formal underpinning for \rmrs's
existing use for partial semantic representations in implemented
systems.  Next to the peace of mind that comes with the use of a
well-understood formalism, we hope that the work reported here will
serve as a starting point for future work on efficient solvers and
inference mechanisms.

@@

Furthermore, the above definition of \rmrs\ provides new concepts
which can help us phrase questions of practical grammar engineering in
well-defined formal terms.  For instance, one crucial issue in
developing a hybrid system that combines or compares the outputs of
deep and shallow processors is to determine whether the \rmrs s
produced by the two systems are compatible.  In the new formal terms
we can use now, we can characterise compatibility of a more detailed
\rmrs\ $\varphi$ (perhaps from a deep grammar) and a less detailed
\rmrs\ $\varphi'$ simply as \emph{entailment} $\varphi \models
\varphi'$.  If entailment holds, this tells us that all claims that
$\varphi'$ makes about the semantic content of a sentence is
consistent with the claims that $\varphi$ makes.

At this point, we cannot provide an efficient algorithm for testing
entailment of \rmrs.  However, we propose the following novel
syntactic characterisation which may again serve as a starting point
for research along those lines. We call an \rmrs\ $\varphi'$ an
\emph{extension} of the \rmrs\ $\varphi$ if $\varphi'$ contains all
the EPs of $\varphi$ and $D(\varphi') \supseteq D(\varphi)$.

>>>>

The above definitions serve an important theoretical purpose in that
they provide a formal underpinning for the use of \rmrs\ in practical
systems.  Next to the peace of mind that comes with the use of a
well-understood formalism, we hope that the work reported here will
serve as a starting point for future research.

@@

Furthermore, the above definition of \rmrs\ provides new concepts
which can help us phrase questions of practical grammar engineering in
well-defined formal terms.  For instance, one crucial issue in
developing a hybrid system that combines or compares the outputs of
deep and shallow processors is to determine whether the \rmrs s
produced by the two systems are compatible.  In the new formal terms,
we can characterise compatibility of a more detailed \rmrs\ $\varphi$
(perhaps from a deep grammar) and a less detailed \rmrs\ $\varphi'$
simply as \emph{entailment} $\varphi \models \varphi'$.  If entailment
holds, this tells us that all claims that $\varphi'$ makes about the
semantic content of a sentence are consistent with the claims that
$\varphi$ makes.

At this point, we cannot provide an efficient algorithm for testing
entailment of \rmrs.  However, we propose the following novel
syntactic characterisation as a starting point for research along
those lines. We call an \rmrs\ $\varphi'$ an \emph{extension} of the
\rmrs\ $\varphi$ if $\varphi'$ contains all the EPs of $\varphi$ and
$D(\varphi') \supseteq D(\varphi)$.

### 2008-10-10 trunk/Papers/rmrs-08/rmrs.tex
final?

Def.\ref{defn:rmrs-syntax}
combines similarities to earlier
presentations of \rmrs\ \cite{copestake:2003,copestake:2007b} and to
{\sc clls}/dominance constraints \cite{egg:etal:2001}.  For the most part,
our syntax generalises that of older versions of \rmrs: We use
$\ARG_{\{i\}}$ (with a singleton set $S$) instead of $\ARG_i$ and
$\ARG_\N$ instead of $\ARG_n$, and the {\sc ep} $\ep{l}{a}{P(v)}$
(as in Section~\ref{sec:motivation}) is an abbreviation of
$\{\ep{l}{a}{P}, \ARG_{\{0\}}(a,v)\}$.  Similarly, we don't assume
that labels, anchors, and holes are syntactically different objects;
they receive their function from their positions in the formula.  One
major difference is that we use dominance rather than qeq; see
Section~\ref{sec:extensions} for a discussion.  Compared to dominance
constraints, the primary difference is that we now have a mechanism
for representing lexical ambiguity, and we can specify a predicate and
its arguments separately.

@@

In this paper, we assume for simplicity that the base language is as
in \mrs; essentially, $\tau$ becomes the structure tree of a formula
of predicate logic.  We assume that $\Sigma$ is a ranked signature
consisting of the symbols of predicate logic: a unary constructor
$\neg$ and binary constructors $\wedge$, $\rightarrow$, etc.; a set of
3-place quantifier symbols such as $\sem{\_every\_q\_1}$ and
$\sem{\_some\_q\_1}$ (with the children being the bound variable, the
restrictor, and the scope); and, constructors of various arities for
the predicate symbols; e.g., $\sem{\_chase\_v\_1}$ is of arity 3.  Other
base languages may require a different signature $\Sigma$ and/or a
different mapping between formulas and trees; the only strict
requirement we make is that the signature contains a binary
constructor $\wedge$ to represent conjunction.  We write $\Sigma_i$
for the set of all constructors in $\Sigma$ with arity $i$.  We
will follow the typographical convention that non-logical symbols in
$\Sigma$ are written in $\mathsf{sans{-}serif}$, as opposed to the
\rmrs\ predicate symbols like $\sempred{\_cat\_n}$ and
$\sempred{\_cat\_n\_1}$.

@@

Because solved forms are so restricted, we can `read off' at least one
model from each solved form:
\begin{prop} \label{prop:solved-forms-are-satisfiable}
  Every {\sc rmrs} in solved form is satisfiable.
\end{prop}
\begin{proof}[Proof (sketch; see also
  \cite{Duchier00dominanceconstraints})] 
  For each {\sc ep}, we choose to label the anchor with the
  constructor $p$ of sufficiently high arity whose existence we
  assumed; we determine the edges between an anchor and its children
  from the uniquely determined $\ARG$ atoms; plugging labels into
  holes is straightforward because no label is dominated by more than
  one label; and  spaces between the labels and
  anchors are filled with conjunctions.
\end{proof}

@@

Let's look at an example for all these definitions.  All the \rmrs s
presented in Section~\ref{sec:motivation} (replacing $\qeq$ by
$\dom$), are in solved form; this is least obvious for
(\ref{ex:cat-erg}), but becomes clear once we notice that no label is
on the right-hand side of two dominance atoms.  However, this is
because (at least) one model of (\ref{ex:cat-erg}) is
shown in Fig.~\ref{fig:fat-black-cat}; this model
is problematic in several ways and in particular contains an unbound
variable $y$ even though it also contains a quantifier that binds $y$.
If we restrict the class of models to those in 
which such variables are bound (as \newcite{copestake:etal:2005} do),
we can add the information that the 
quantifiers outscope their bound variables to the \rmrs\ without
changing its models further---i.e., we add the atoms $h_3 \dom
l_5$ and $h_8 \dom l_5$.  Fig.~\ref{fig:fat-black-cat} is no longer a
model for the extended \rmrs, which in turn is no longer in solved
form because the label $l_5$ is on the right-hand side of two
dominance edges.  Instead, it has the following two solved forms:

@@

Furthermore, grammar developers such as those in the {\sc delph-in}
community typically adopt conventions that restrict them to a fragment
of the language from Def.~\ref{defn:rmrs-syntax} (once qeq is added to
it), or they restrict attention to only a subset of the models.
Our formalism provides a
general framework into which all these various fragments fit, and it's
a matter of future work to explore these fragments further.

>>>>

Def.~\ref{defn:rmrs-syntax} combines similarities to earlier
presentations of \rmrs\ \cite{copestake:2003,copestake:2007b} and to
{\sc clls}/dominance constraints \cite{egg:etal:2001}.  For the most
part, our syntax generalises that of older versions of \rmrs: We use
$\ARG_{\{i\}}$ (with a singleton set $S$) instead of $\ARG_i$ and
$\ARG_\N$ instead of $\ARG_n$, and the {\sc ep} $\ep{l}{a}{P(v)}$ (as
in Section~\ref{sec:motivation}) is an abbreviation of
$\{\ep{l}{a}{P}, \ARG_{\{0\}}(a,v)\}$.  Similarly, we don't assume
that labels, anchors, and holes are syntactically different objects;
they receive their function from their positions in the formula.  One
major difference is that we use dominance rather than qeq; see
Section~\ref{sec:extensions} for a discussion.  Compared to dominance
constraints, the primary difference is that we now have a mechanism
for representing lexical ambiguity, and we can specify a predicate and
its arguments separately.

@@

In this paper, we assume for simplicity that the base language is as
in \mrs; essentially, $\tau$ becomes the structure tree of a formula
of predicate logic.  We assume that $\Sigma$ is a ranked signature
consisting of the symbols of predicate logic: a unary constructor
$\neg$ and binary constructors $\wedge$, $\rightarrow$, etc.; a set of
3-place quantifier symbols such as $\sem{\_every\_q\_1}$ and
$\sem{\_some\_q\_1}$ (with the children being the bound variable, the
restrictor, and the scope); and constructors of various arities for
the predicate symbols; e.g., $\sem{\_chase\_v\_1}$ is of arity 3.  Other
base languages may require a different signature $\Sigma$ and/or a
different mapping between formulas and trees; the only strict
requirement we make is that the signature contains a binary
constructor $\wedge$ to represent conjunction.  We write $\Sigma_i$
for the set of all constructors in $\Sigma$ with arity $i$.  We
will follow the typographical convention that non-logical symbols in
$\Sigma$ are written in $\mathsf{sans{-}serif}$, as opposed to the
\rmrs\ predicate symbols like $\sempred{\_cat\_n}$ and
$\sempred{\_cat\_n\_1}$.

@@

Because solved forms are so restricted, we can `read off' at least one
model from each solved form:
\begin{prop} \label{prop:solved-forms-are-satisfiable}
  Every {\sc rmrs} in solved form is satisfiable.
\end{prop}
\begin{proof}[Proof (sketch; see also
  \cite{Duchier00dominanceconstraints})] 
  For each {\sc ep}, we choose to label the anchor with the
  constructor $p$ of sufficiently high arity whose existence we
  assumed; we determine the edges between an anchor and its children
  from the uniquely determined $\ARG$ atoms; plugging labels into
  holes is straightforward because no label is dominated by more than
  one hole; and spaces between the labels and anchors are filled with
  conjunctions.
\end{proof}

@@

Let's look at an example for all these definitions.  All the \rmrs s
presented in Section~\ref{sec:motivation} (replacing $\qeq$ by
$\dom$), are in solved form; this is least obvious for
(\ref{ex:cat-erg}), but becomes clear once we notice that no label is
on the right-hand side of two dominance atoms.  However, the model
constructed in the proof of
Prop.~\ref{prop:solved-forms-are-satisfiable} looks a bit like
Fig.~\ref{fig:fat-black-cat}; both models are problematic in several
ways and in particular contain an unbound variable $y$ even though
they also contains a quantifier that binds $y$.  If we restrict the
class of models to those in which such variables are bound (as
\newcite{copestake:etal:2005} do), we can enforce that the quantifiers
outscope their bound variables without changing models of the \rmrs\
further---i.e., we add the atoms $h_3 \dom l_5$ and $h_8 \dom l_5$.
Fig.~\ref{fig:fat-black-cat} is no longer a model for the extended
\rmrs, which in turn is no longer in solved form because the label
$l_5$ is on the right-hand side of two dominance atoms.  Instead, it
has the following two solved forms:

@@

Furthermore, grammar developers such as those in the {\sc delph-in}
community typically adopt conventions that restrict them to a fragment
of the language from Def.~\ref{defn:rmrs-syntax} (once qeq is added to
it), or they restrict attention to only a subset of the models (e.g.,
ones with correctly bound variables, or ones which don't contain extra
material like Fig.~\ref{fig:fat-black-cat}).  Our formalism provides a
general framework into which all these various fragments fit, and it's
a matter of future work to explore these fragments further.
